---
title: "P8130 Biostatistical Methods I - Homework 1"
date: "22 September 2024"
output: github_document
---

```{r echo = FALSE, message = FALSE}
library(tidyverse)
```

# Question 2
## Bike Crash History
The following data set represents 14 individuals with recent bike crash history: 

45, 39, 25, 47, 49, 5, 70, 99, 74, 37, 99, 35, 8, 59

The following represents the mean, median, range, and standard deviation of the depression scores, as well as a summary of the key characteristics.

```{r}
# bike crash history depression scores and key characteristics
bike = c(45, 39, 25, 47, 49, 5, 70, 99, 74, 37, 99, 35, 8, 59)

mean(bike)
median(bike)
max(bike) - min (bike)
sd(bike)
summary(bike)
```


## Car Crash History
The following data set represents 13 individuals with recent car crash history:

67, 50, 85, 43, 64, 35, 47, 97, 58, 58, 10, 56, 50

The following represents the mean, median, range, and standard deviation of the depression scores, as well as a summary of the key characteristics.
```{r}
# car crash depression scores dataset and key characteristics
car = c(67, 50, 85, 43, 64, 35, 47, 97, 58, 58, 10, 56, 50)

mean(car)
median(car)
max(car) - min(car)
sd(car)
summary(car)
```

## Combining Two Vectors Using cbind
```{r, warning = FALSE}
# combining the two vectors of bike crash history depression scores and car crash history depression scores
crash_history = cbind(bike, car)
```

## Pass crash_history to a Boxplot
```{r}
# boxplot of depression scores stratified by type of accident
boxplot(
  crash_history, 
  title = "Boxplot of Depression Scores by Type of Accident", 
  ylab = "Depression Scores", 
  xlab = "Type of Accident",
  col = c("lavender", "lemonchiffon") 
)
```

# Question 4
## Given Probabilities
```{r}
# probability of having dementia
P_D = 0.05

# probability of positive CT scan given she has dementia
P_P_given_D = 0.80

# probability of positive CT scan given she does not have dementia
P_P_given_no_D = 0.10
```

## Calculating P
```{r}
# probability of not having dementia
P_no_D = 1 - P_D
```

## Law of Total Probability to calculate P(P)
```{r}
# probability of positive CT scan
P_P = (P_P_given_D * P_D) + (P_P_given_no_D * P_no_D)
```

## Bayes Theorem
```{r}
# probability of having dementia given a positive CT scan
P_D_given_P = (P_P_given_D * P_D)/(P_P)
```

